---
title: "Activity Prediction"
author: "Vipin"
date: "4/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, we will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

For the purpose of this project, the following steps would be followed:

1. Data Preprocessing
2. Exploratory Analysis
3. Prediction Model Selection
4. Predicting Test Set Output

## Data Preprocessing 
Loading data
```{r cache=TRUE}
library(caret)
trainURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainURL))
testing <- read.csv(url(testURL))
label <- createDataPartition(training$classe, p = 0.7, list = FALSE)
train <- training[label, ]
test <- training[-label, ]
dim(train)
```

Cleaning Data: Data contains 160 variable, out of which many have very less variance, some have lots of NA and some are only for identification.
```{r cache=TRUE}
NZV <- nearZeroVar(train)
train <- train[ ,-NZV]
test <- test[ ,-NZV]
label <- apply(train, 2, function(x) mean(is.na(x))) > 0.95
train <- train[, -which(label, label == FALSE)]
test <- test[, -which(label, label == FALSE)]
train <- train[ , -(1:5)]
test <- test[ , -(1:5)]
dim(train)
```
So, now we are left with only 54 variables.

## Exploratory Analysis
```{r fig.width=12, fig.height=8}
library(corrplot)
corrMat <- cor(train[,-54])
corrplot(corrMat, method = "color", type = "lower", tl.cex = 0.8, tl.col = rgb(0,0,0))
```

So there are few correlations in the data. PCA can be use to further remove these correlations. But, as these correlations are very less, so we won't be applying it.

### Principal component analysis


## Prediction Model Selection
We will use 3 methods to model the training set and thereby choose the one having the best accuracy to predict the outcome variable in the testing set.   
The methods are 
1. Decision Tree  
2. Random Forest  
3. Generalized Boosted Model  

### Decision Tree 
```{r fig.width=18, fig.height=20, cache=TRUE}
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(1590)
library(caret)
modelDT <- rpart(classe ~ ., data = train, method = "class")
fancyRpartPlot(modelDT)
predictDT <- predict(modelDT, test, type = "class")
confMatDT <- confusionMatrix(predictDT, test$classe)
confMatDT
```

### Random Forest

```{r cache=TRUE}
library(caret)
set.seed(1590)
control <- trainControl(method = "cv", number = 3, verboseIter=FALSE)
modelRF <- train(classe ~ ., data = train, method = "rf", trControl = control)
modelRF$finalModel
predictRF <- predict(modelRF, test)
confMatRF <- confusionMatrix(predictRF, test$classe)
confMatRF
```

### Generalized Boosted Model

```{r cache=TRUE}
library(caret)
set.seed(1590)
modelGBM <- train(classe ~ ., data = train, method = "gbm", verbose = FALSE)
modelGBM$finalModel
control <- trainControl(method = "repeatedcv", number = 5, repeats = 1, verboseIter = FALSE)
modelGBM <- train(classe ~ ., data = train, trControl = control, method = "gbm", verbose = FALSE)
confMatGBM <- confusionMatrix(predictGBM, test$classe)
confMatGBM
```

As Random Forest offers the maximum accuracy, we will go with Random Forest Model to predict our test data class variable.

## Predicting Test Set Output

```{r}
predictRF <- predict(modelRF, testing)
predictRF
```


